---
title: "NFL Lower Limb, Cleaning Up"
author: "Parker Pontarolo"
date: "2023-01-19"
output: slidy_presentation
---
## Before I Start

I would Provide the data but the "PlayerTrackData.csv" is too big to share on Github

=================================================TIPS=================================================
when you see 
```{r}
```
in RStudio I am using a Markdown file. This makes a R script between those three punctuation.

What is between those punctuation are being compiled, everything outside of it, is not and are notes for my Slidy for RPubs
=======================================================================================================

## Data

https://www.kaggle.com/competitions/nfl-playing-surface-analytics/data

```{r}

## if possible run this chunk ONCE
Injury = read.csv("InjuryRecord.csv")
Player = read.csv("PlayerTrackData.csv")
Play = read.csv("PLayList.csv")

```
```{r}
library(tidyverse)
library(ggplot2)
```

## Cleaning up Data

```{r}

head(Injury)

Injury_Clean = Injury %>%
            rename("1_Day_or_More" = "DM_M1",
                   "7_Days_or_More" = "DM_M7",
                   "4_Weeks_or_More" = "DM_M28",
                   "6_Weeks_or_More" = "DM_M42")

head(Injury_Clean)

```

```{r}

head(Player)

Player_Clean = Player %>% 
            rename("Direction" = "dir",
                   "Distance" = "dis",
                   "Orientation" = "o",
                   "Yards_per_Second" = "s",
                   "Seconds" = "time",
                   "Play_Detail" = "event")

head(Player_Clean)

```
```{r}

head(Play)

```


After looking at the Type of Stadiums in Excel. I've noticed something weird, Alot of Outdoors and Indoors. It is what i expected, but i caught something that you can see here..
```{r}

Play$StadiumType

```
outdoors, outdoor, oudoor.

## Lets fix that



```{r}

Outdoor = c("Oudoor", "Ourdoor", "Outddors", "Outdoor", "Outdoor Retr Roof-Open", "Outdoors", "Outdor", "Outside", "Heinz Field", "Cloudy", "Bowl")
Indoor_Closed_Roof = c("Indoor", "Indoor, Roof Closed", "Retr. Roof-Closed", "Retr. Roof - Closed", "Retr. Roof Closed", "Retractable Roof", "Indoors", 
                       "Domed, closed", "Domed", "Dome, closed", "Dome", "Closed Dome")
Indoor_Open_Roof = c("Retr. Roof-Open", "Retr. Roof - Open", "Indoor, Open Roof", "Open", "Domed, Open", "Domed, open")

Stadium_Clean = function(x)
                {
                  if(x %in% Outdoor)
                  {
                    "Outdoor"
                  }
                  else if(x %in% Indoor_Closed_Roof)
                  {
                    "Indoor_Closed_Roof"
                  }
                  else if(x %in% Indoor_Open_Roof)
                  {
                    "Indoor_Open_Roof"
                  }
                  else
                  {
                    "Not_Recorded"
                  }
                }

Play_Clean = Play %>%
              mutate(StadiumType = mapply(Stadium_Clean, StadiumType))

Play_Clean %>% 
    count(StadiumType)
```
## Trust Issues

Now i'm worried that others are like this...
Here are the ones I checked 
```{r}

Play %>% 
    count(StadiumType)
  
```
```{r}
Injury%>%
  count(BodyPart)

```
```{r}

Injury%>%
  count(Surface)

```
```{r}

Player_Clean %>%
  count(Play_Detail)

```
```{r}

Play_Clean %>%
  count(RosterPosition)

```
```{r}
Play_Clean %>%
  count(Weather)
```

## Found another one

```{r}

Rain = c("10% Chance of Rain", "30% Chance of Rain", "Cloudy, 50% change of rain", "Rain Chance 40%", "Rain likely, temps in low 40s.", "Rain", "Rain shower", "Rainy", "Showers", "Scattered Showers", "Light Rain", "Cloudy, Rain", "Cloudy with periods of rain, thunder possible. Winds shifting to WNW, 10-20 mph.", "Cloudy, chance of rain")

Cloudy = c("cloudy", "Cloudy", "Overcast","Mostly cloudy", "Mostly Cloudy", "Mostly Coudy", "Sun & clouds", "Party Cloudy", "Partly Clouidy", "Partly Cloudy", "Partly cloudy", "Coudy", "Clear to Partly Cloudy", "Cloudy, fog started developing in 2nd quarter", "Hazy")

Snow = c("Snow", "Heavy lake effect snow", "Cloudy, light snow accumulating 1-3 inches")

Cold = c("Cloudy and cold", "Cloudy and Cool", "Cold", "Sunny and cold", "Clear and cold", "Clear and Cool")

Indoor = c("N/A (Indoors)", "N/A Indoor", "Indoors", "Indoor", "Controlled Climate")

Clear = c("Clear", "Clear and sunny", "Clear and Sunny", "Clear and warm", "Clear skies", "Clear Skies", "Fair", "Partly clear", "Sunny and clear")

Sunny = c("Heat Index 95", "Mostly sunny", "Mostly Sunny", "Mostly Sunny Skies", "Partly sunny", "Partly Sunny", "Sunny", "Sunny and warm", "Sunny Skies", "Sunny, highs to upper 80s", "Sunny, Windy")

Weather_Clean = function(x)
                {
                  if(x %in% Rain)
                  {
                    "Rain"
                  }
                  else if(x %in% Cloudy)
                  {
                    "Cloudy"
                  }
                  else if(x %in% Snow)
                  {
                    "Snow"
                  }
                  else if(x %in% Cold)
                  {
                    "Cold"
                  }
                  else if(x %in% Indoor)
                  {
                    "Indoor"
                  }
                  else if(x %in% Clear)
                  {
                    "Clear"
                  }
                  else if(x %in% Sunny)
                  {
                    "Sunny"
                  }
                  else
                  {
                    "Not_Recorded" # this is the cell that was blank
                  }
                }

Play_Clean = Play_Clean %>%
              mutate(Weather = mapply(Weather_Clean, Weather))

Play_Clean %>% 
    count(Weather)


```
## Save

Save the data after cleaning them
```{r}

## comment out the saves because i already did that

##save(Play_Clean, file = "Play.RData")
##save(Injury_Clean, file = "Injury.RData")
##save(Player_Clean, file = "Player.RData")

```

